Setup Honeypot: Team Contagiarium

### 2011-11-01 Setup_honeypot.tex
Typo
More sections

After these period we will analyze the results of both setups and compare the differences in the results we have captured. 
The advantage of a LI honeypot is that the machine can only be compromised to a certain extent and can not really be used for any malicious activity.
The disadvantage is that it is not possible to capture all the activity an attacker might perform, like after escalating privileges - which should not be possible.
The advantage of a HI honeypot is that you can capture all the activity, but that the system might also be completely compromised and do damage to other systems.
That is why we need to be much more vigilant in monitoring our honeypot in the HI phase.

@@

\subsection{Syslog}
By default, \verb|syslog| uses UDP to transmit its log messages to a remote server.
Relying on UDP is not very desirable, because everything is sent in plaintext and delivery of UDP messages is not guaranteed.
To allow for safe and reliable transmission of the syslog messages from the honeypot to the monitoring machine, one needs to use TCP connections in syslog, and provide some way of encapsulating the socket used in a cryptographic layer.

%A VPN session between the two hosts is possible, but might be considered as overkill and still does not garantee delivery.
We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with SSH and syslog-ng''. \cite{remote_logging}
The example command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.
We have put this command in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason.

We have configured a locally installed the \verb|syslog-ng| package for our user on the monitoring machine, and configured it to listen to local TCP connections on port 9514 and thereby receive syslog messages transmitted through this port.

@@

@@

In this section we decribe the setup of the honeypot.

\subsection{Rsyslog}
On the honeypot, we've modified the \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain.
The modified configuration is now stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking interest.
The configuration should and will be moved over to one of the default configuration files to avoid raising suspicion.
Note that the chosen port is very similar to the default syslog port 514, and this might give us away.
We plan to change it to a more nondescript port number.

@@

\subsection{Kippo}
To capture attackers that try to come in by brute-forcing SSH logins, we are using Kippo. 
Kippo is used to log bruteforce attacks and the shell interaction performed by the attacker. 
To make this as realistic as possible we have added a redirect in the iptables configuration from port 22 to port 2222 that kippo uses. 
Unfortunately this broke down around the time the firewall was being reconfigured, for unclear reasons. 
The reason for forwarding was not to give Kippo root capabilities.

@@

@@

To set up GHH we have first installed Apache and PHP5.
We configured GHH to log to in CSV format, since it requires minimal setup and is easily parsed.
%We can later reconfigure to use MySQL, if the results are not satisfactory.
Then, we patched the GHH code to allow logging to syslog, to maintain a single, consistent point of log collection.

@@

We will configure Snort to send alerts to syslog, and to log its packet captures to a read-only \verb|sshfs| mounted from our monitoring machine.
We will install a script like \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.

>>>>

After these periods we will analyze the results of both setups and compare the differences in the results we have captured. 
The advantage of a LI honeypot is that the machine can only be compromised to a certain extent and can not really be used for any malicious activity.
The disadvantage is that it is not possible to capture all the activity an attacker might perform, like after escalating privileges - which should not be possible.
The advantage of a HI honeypot is that you can capture all the activity, but that the system might also be completely compromised and do damage to other systems.
That is why we need to be much more vigilant in monitoring our honeypot in the HI phase.

@@

\subsection{Syslog-ng}
By default, \verb|syslog| uses UDP to transmit its log messages to a remote server.
Relying on UDP is not desirable, because everything is sent in plaintext and delivery of UDP messages is not guaranteed.
To allow for confidential and reliable transmission of the syslog messages from the one machine to another, one needs to use TCP connections in syslog, and provide some way of encapsulating the socket used in a cryptographic layer.

%A VPN session between the two hosts is possible, but might be considered as overkill and still does not garantee delivery.
We have decided to set up a reverse SSH tunnel from the monitoring machine to the honeypot machine, as described in the article ``Remote logging with SSH and syslog-ng''. \cite{remote_logging}
The example command establishes an SSH connection the honeypot, setting options to forego a login shell or anything a normal user would desire, and binds the local socket 9514 on the honeypot machine to tunnel its connections to the same port on our monitoring machine.

We have configured a locally installed the \verb|syslog-ng| package for our user on the monitoring machine, have configured it to listen to local TCP connections on port 9514 and will thereby receive syslog messages transmitted through this port.
We are aware there is still the possibility of other users on the monitoring machine poisoning our logs, but we feel we can rule out foul play by our peers in this assignment.

@@

We have put the command to create the SSH tunnel in a script to allow the connection to be re-esthablished quickly and easily if the SSH session is terminated for some reason. 
A severance of the connection would lead us to assume the worst initially.
If re-establishing this connection fails after any attackers have gained access (which should never happen if we monitor adequately), we will have to assume the machine has been compromised beyond our control and will request a shutdown of the VM, if no network downtime was otherwise announced.

@@

In this section we describe the setup of the honeypot.

\subsection{Rsyslog}
On the honeypot, we have modified the default \verb|rsyslog| configuration to also send syslog messages to the (now bound) local port 9514, completing the chain to the syslog instance on our monitoring machine.
The modified configuration is now stored in its own file in the syslog.d directory, which is a bit obvious to anyone taking an interest.
We plan to move this small configuration over to one of the default configuration files to avoid raising suspicion outright.
Note that the chosen port for remote syslog is very similar to the default syslog port 514, and this might give us away.
We plan to change it to a more nondescript port number.

\subsection{SSHFS mount}
Not all services support logging to syslog, and for some it is quite infeasible.
We will create a read-only \verb|sshfs| mount from our monitoring machine that we will use to write our logfiles for various services to.
We will use interactive login to mount this, as allowing the honeypot back into our monitoring machine with an SSH keypair just seems dangerous.

@@

\subsection{Kippo}
To capture attackers that try to come in by brute-forcing SSH logins, we are using Kippo. 
Kippo is used to log bruteforce attacks and the shell interaction performed by the attacker. 
We've installed the latest development version to run as a separate user.
To make this as realistic as possible we have added a redirect in the iptables configuration from port 22 to port 2222 that kippo uses by default.
Unfortunately this broke down around the time the firewall was being reconfigured, for unclear reasons. 
The reason for forwarding was not to give Kippo root capabilities.

@@

This does not log to syslog. We hope to fix that, but it's not a priority until the HI phase.

\section{Apache \& Nginx}
To attract some web traffic, we have installed the Apache web server, the Apache PHP runtime and the \verb|mod-security| module which allows us to spoof its version information, to make it appear old and vulnerable.
The Nginx webserver is installed as a reverse proxy in front of Apache, which runs on local port 8080.
This provides 2 layers of logging (to files), and Nginx's superior transparent caching mechanism allows us to avoid exhausting our resources with scripting languages, should we ever face an application-layer DoS attack (Nginx appears quite resilient against those).

Nginx has been compiled from source and was patched to send the spoofed Apache version header as well.
We will configure both services to log to our \verb|sshfs| mount.

@@

%To set up GHH we have first installed Apache and PHP5.
We configured GHH to log to in CSV format, since it requires minimal setup and is easily parsed.
%We can later reconfigure to use MySQL, if the results are not satisfactory.
Then, we patched the GHH code to allow logging to syslog, to maintain a single, consistent point of log collection.

@@

We will configure Snort to send alerts to syslog, and to log its packet captures to the read-only \verb|sshfs| mounted from our monitoring machine.
We will install a script like \verb|oinkmaster| to automatically retrieve the latest community Snort rules, with a Snort API key registered specifically for this project.
